



<head>
  <meta charset="utf-8" />
  <title>TB Daily Status</title>
  <link rel="stylesheet" type="text/css" href="../css/jquery-ui.css" />
  <script src="http://code.jquery.com/jquery-1.9.1.js"></script>
  <script src="http://code.jquery.com/ui/1.10.3/jquery-ui.js"></script>
  <link rel="stylesheet" type="text/css" href="../css/tb_report.css" />
  <script>
  $(function() {
    $( "#tabs" ).tabs();
  });
  </script>
</head>


```{r setup, echo= FALSE, message=FALSE,warning= FALSE}

# Strings ain't factors
options(stringsAsFactors = FALSE)

# Load the required libraries
library(knitr)
library(mpmisc)
library(RODBC)
library(plyr)
library(ggplot2)
library(scales)
library(gridExtra)
library(lubridate)
library(reshape2)



# Set default chunk options
opts_chunk$set(echo = FALSE,
               results = 'asis',
               message = FALSE,
               warning = FALSE,
               fig.width = 10,
               fig.cap = "",
               error = TRUE)




# Set period dates
# Go twelve weeks back from the next Monday - this helps ensure that
# plotted weeks are full weeks. Otherwise, you might end up with just a Friday,
# which makes a given week look really slow.
# Obviously, the week in progress will always appear as a partial week.

start_date <- get_week_start(Sys.Date(), weeks.back = 11)
stop_date <- Sys.Date()


# Set some default plot parameters
point_size <- 3
line_size <- 1.3
colorwheel <- c("#31A354", "#ADDD8E", "#F03B20", "#43A2CA")

theme_tb <- theme_bw() +
            theme(legend.key.width = unit(.75, "inches"))



# Set up the database connection
plus <- connect_to_tbdbplus()



```






# TB Clinic Daily Status Report
Updated at `r paste(Sys.time())`


<div id="tabs">
  <ul>
    <li><a href="#alerts">Alerts</a></li>
    <li><a href="#cases">Cases</a></li>
    <li><a href="#visits">Visits</a></li>
    <li><a href="#diagnostics">Diagnostics</a></li>
    <li><a href="#treatment">Treatment</a></li>
    <li><a href="#contacts">Contacts</a></li>
  </ul>
    


  <div id="alerts">

## Patients Requiring 60-Day Sputa

```{r sputa}

# NTIP criteria: TB patients with positive sputum culture results alive at diagnosis and have
# initiated treatment, counted in the year of interest. Patients who died within 60 days of
# initiating treatment are excluded. 


# Query the RVCT, using TBdb's mycobacteriology records to identify the culture-positive
rvct <- sqlQuery(plus, "
                 
    SELECT r.person_id, p.last_name, p.first_name, p.date_of_birth, r.therapystart
    FROM RVCT r LEFT OUTER JOIN Person p
    ON r.person_id = p.idnumber
    WHERE r.therapystart >= (Date() - 90)
        AND r.alive = 1
        AND r.reportingcountyid IN ('01', '03', '07', '08', '09', '17', '19')
        AND (r.diseasesite1 = 1 OR r.diseasesite2 = 1 OR r.diseasesite3 = 1)
        AND r.person_id IN (
            SELECT person_id
            FROM Culture
            WHERE Culture_Status = '12'
        )
    ORDER BY r.therapystart
                 
")




# Get their sputa results
sputa <- sqlQuery(plus, paste("
                  
    SELECT person_id, collection_date, culture_status
    FROM Culture
    WHERE collection_date >= (Date() - 90)
        AND specimen_source = '1'
        AND person_id IN (",
        paste(rvct$person_id, collapse = ", "),
        ")",
    sep = "")

)



# Mark the 52 to 60-day window for sputum collections
rvct$sputwin52 <- as.Date(rvct$therapystart) + 52
rvct$sputwin60 <- as.Date(rvct$therapystart) + 60



# Identify whether patients have a sputum collected during that window
rvctsput <- ddply(rvct, .var = "person_id", .fun = function(x) {
    
    # Do any of their documented sputa fall in the window?
    xsput <- subset(sputa,
                    subset = person_id %in% x$person_id &
                             as.Date(collection_date) >= x$sputwin52 &
                             as.Date(collection_date) <= x$sputwin60
    )
    
    # Flag if a sputum was collected during the window, and if all sputa collected were negative
    mutate(x,
           anywindowsput = nrow(xsput) > 0,
           windowsputneg = sum(xsput$culture_status %in% 2) %in% nrow(xsput) & nrow(xsput) > 0
    )
    
})
    
    




# Recode nwindowsput and windowsputneg
rvctsput$anywindowsput[rvctsput$anywindowsput %in% FALSE] <- "No"
rvctsput$anywindowsput[rvctsput$anywindowsput %in% TRUE] <- "Yes"

rvctsput$windowsputneg[rvctsput$windowsputneg %in% FALSE] <- "No"
rvctsput$windowsputneg[rvctsput$windowsputneg %in% TRUE] <- "Yes"





# Pretty names
names(rvctsput) <- c("TBdb ID", "Last Name", "First Name", "DOB", "Tx Start", 
                     "Sputum Window: 52 Days", "Sputum Window: 60 Days",
                     "Sputum Collected During Window?", "Window Sputa All Negative?")

dfprintif(rvctsput,
          norowmsg = "All culture-positive patients appear to have sputum conversions documented.")



```


## Missed LTBI Pickups in the Last 14 Days

  ```{r missed_pickups}
  
  # Query the missed LTBI pickups from the last fourteen days
  missed_pu <- sqlQuery(plus, paste("
                                    
      SELECT t.person_id, p.last_name, p.first_name, t.treatment_date
      FROM Drug_Treatment t LEFT OUTER JOIN Person p
      ON t.person_id = p.idnumber
      WHERE dispense_type = 'Pickup'
          AND completed = ''
          AND treatment_date BETWEEN (Now() - 14) AND (Now() - 1)
      ORDER BY treatment_date",
                                    
      sep = "")
                        
  )
  
  
  # Add treatment plan author
  # Get each individual's most recent treatment plan
  missed_plans <- sqlQuery(plus, "
  
    SELECT person_id, plan_author, author_affiliation, treat_plan_date
    FROM Tx_Plan_View
    ORDER BY person_id, treat_plan_date DESC
    
  ")
  
  latest_author <- missed_plans[!duplicated(missed_plans$person_id), ]
  
  
  # Merge the latest plan author onto the treatments
  missed_author <- merge(x = missed_pu,
                         y = latest_author,
                         by = "person_id",
                         all.x = TRUE)
                         
  # Print out only the DPH plans
  missed_dph <- missed_author[missed_author$author_affiliation %in% "Denver Metro TB Clinic",
                              c("person_id", "last_name", "first_name", 
                                "treatment_date", "plan_author")]
                                    
  
  
  dfprintif(arrange(missed_dph, plan_author, treatment_date),
            norowmsg = "No LTBI pickups have been missed.")
  
  
  ```




  </div>




## Active Cases Over the Last Two Years


  <div id="cases">
  
  
```{r cases}
source("../Quarterly Strategic Area Review/R/fun/query_actives.r")


# Get active cases over the last two years.
# Ensure that the oldest year is a full year
actives <- query_actives(start_date = as.Date(paste(as.numeric(format(Sys.Date(), "%Y")) - 2,
                                                    "-01-01", 
                                                    sep = "")),
                         stop_date = Sys.Date(),
                         odbc = plus)

metroactives <- subset(actives, metro_case %in% 'True')


# Create a quarter label
metroactives$plot_qtr <- with(metroactives, paste(yr_id, " Q", qtr_id, sep = ""))

# Create plot labels for county
metroactives$plot_group <- NA
metroactives$plot_group[metroactives$report_county %in% "Denver"] <- "Denver"
metroactives$plot_group[!metroactives$report_county %in% "Denver"] <- "Other Metro Counties"


# Aggregate into two years of quarterly counts and ten years of annual counts
actives_qtr <- count(metroactives, "plot_qtr")



ggplot(actives_qtr, aes(x = plot_qtr, y = freq, group = 1)) +
    geom_point(size = point_size, color = "#31A354") +
    geom_line(size = line_size, color = "#31A354") +
    expand_limits(y = 0) +
    labs(x = "Quarter identified (earliest of date treatment started or date case reported)", 
         y = "Number of cases",
         title = "Number of Active Cases Identified") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))


```

  
```{r cases_table,results='asis'}


actives_print <- ddply(metroactives, .var = "plot_qtr", .fun = summarise,
                       total = length(report_county),
                       denver = sum(report_county %in% 'Denver'),
                       metro = total - denver,
                       denverprop = as.integer(denver / total * 100)
)
                       

# Print it with nice names
names(actives_print) <- c("Year", "Total Cases", "Denver", "Other Metro Counties",
                          "% Denver")

dfprintif(actives_print)


```
  
  
  
  </div>


  <div id="visits">

## Recent Clinic Visits

```{r visits}


source("../Quarterly Strategic Area Review//R//fun/query_visits.r")

visits <- query_visits(start_date = start_date,
                       stop_date = stop_date,
                       odbc = plus)

# Add a week indicator
visits$visit_week <- format(visits$visit_date, format = "%Y-%W")

# Aggregate by week
locagg <- ddply(subset(visits, location %in% c("Clinic", "Outreach")),
                .var = c("location", "visit_week"),
                .fun = summarise,
                n_visits = length(visit_date),
                week_date = get_week_start(as.Date(max(visit_date, na.rm = TRUE)))
)


totals <- ddply(subset(visits, location %in% c("Clinic", "Outreach")),
                .var = "visit_week",
                .fun = summarise,
                n_visits = length(visit_date),
                week_date = get_week_start(as.Date(max(visit_date, na.rm = TRUE)))
)


totals$location <- "All Visits"

visitagg <- rbind(locagg, totals)



ggplot(visitagg, aes(x = week_date, y = n_visits, group = location, color = location)) +
    geom_point(size = point_size) +
    geom_line(aes(linetype = location), size = line_size) +
    expand_limits(y = 0) +
    scale_x_date(breaks = date_breaks("1 week")) +
    scale_color_manual("Visit Location", values = colorwheel[1:3]) +
    scale_linetype_discrete("Visit Location") +
    labs(x = "Visit Week", 
         y = "Number of visits",
         title = "Patient Visits by Location") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, hjust = 1.30, vjust = 1.3))



```


```{r, results='asis'}

# Cast visits wide
visits_wide <- dcast(visitagg, week_date ~ location, value.var = "n_visits", fun.agg = sum)

# Pretty-print
names(visits_wide) <- c("Visit Week", "Total Visits", "Clinic Visits", "Outreach Visits")

dfprintif(visits_wide)

```

  </div>


  <div id="diagnostics">

## Recent Screenings
```{r screens}

# Load the function
source("../Quarterly Strategic Area Review//R//fun/query_tests.r")

# Query test results for the period
tests <- query_tests(start_date = start_date,
                     stop_date = stop_date,
                       odbc = plus)


# Add a week indicator
tests$test_week <- format(tests$test_date, format = "%Y-%W")


# Aggregate by week
testagg <- ddply(tests,
                .var = c("test", "test_week"),
                .fun = summarise,
                freq = length(test_date),
                week_date = get_week_start(as.Date(max(test_date, na.rm = TRUE)))
)


# Plot it
ggplot(testagg, aes(x = week_date, y = freq, group = test, color = test)) +
    geom_point(size = point_size) +
    geom_line(aes(linetype = test), size = line_size) +
    expand_limits(y = 0) +
    scale_x_date(breaks = date_breaks("1 week")) +
    scale_color_manual("Diagnostic", values = colorwheel[1:3]) +
    scale_linetype_discrete("Diagnostic") +
    labs(x = "Diagnostic date", 
         y = "Number of diagnostics",
         title = "Screenings for Active and Latent Tuberculosis") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, hjust = 1.30, vjust = 1.3))


```


```{r, results='asis'}

# Cast tests wide
tests_wide <- dcast(testagg, week_date ~ test, value.var = "freq")

# Pretty-print
names(tests_wide) <- c("Test Week", "CXRs", "QFTs", "TSTs")

dfprintif(tests_wide)

```




## CXRs Awaiting Action
```{r action_cxrs}

# Query the CXRs with action required from TBdb
action_cxrs <- sqlQuery(plus, "
    
    SELECT x.person_id, x.cxr_date_taken, x.cxr_date_read,
           x.cxr_action_text, x.action_required
    FROM CXR_View x LEFT OUTER JOIN TB_Case c
    ON x.person_id = c.person_id
    WHERE c.agency_id = 30
        AND (x.action_required = 'Yes' OR x.action_required Is Null)
    ORDER BY x.cxr_action_text, x.cxr_date_taken
                         
")

# Convert cxr_date_taken to Date
action_cxrs$cxr_date_taken <- as.Date(action_cxrs$cxr_date_taken)

# There are a lot of CXRs with action_required = NA; we only need to address
# those in the last year
action_cxrs_recent <- subset(action_cxrs,
                             subset = action_required %in% "Yes" |
                                      (is.na(action_cxrs$action_required) & 
                                      action_cxrs$cxr_date_taken >= (Sys.Date() - 365)),
                             select = !names(action_cxrs) %in% "action_required"
)




dfprintif(action_cxrs_recent, 
          norowmsg = "It appears that there are no upcoming LTBI pickups at this time.")

```

  </div>


<div id="treatment">


## Patients Receiving Treatment
```{r tx_plot}

# Query all plans
plans <- sqlQuery(plus, "
                  
    SELECT t.person_id, t.treat_plan_type,
           t.treat_plan_date, t.treat_plan_end
    FROM Tx_Plan_View t LEFT OUTER JOIN Demos_view d
    ON t.person_id = d.person_id
    WHERE author_affiliation = 'Denver Metro TB Clinic'
        AND treat_plan_date > #2009-01-01#
                      
")


# Over the last twelve weeks, how many people were on treatment?
# For each week, I need to determine if a given treatment plan started before or during the week
# and if it ended during or after the week
# Get the Fridays of interest
weeks <- data.frame(friday = get_week_start(Sys.Date(), start.day = "Friday", weeks.back = 0:10))

# Get their corresponding Mondays
weeks <- ddply(weeks, .var = "friday", .fun = summarise,
    monday = get_week_start(friday, start.day = "Monday")
)
                    

weeklytx <- ddply(weeks, .var = "monday", .fun = function(x) {
    
    xplans <- plans
    
    # Tag plans that started before or during the week, 
    # and those that ended during or after (or are still in progress, of course)
    xplans$start_bod <- with(xplans, 
                             as.Date(treat_plan_date) <= x$friday
    )
    
    xplans$end_doa <- with(xplans, 
                           (as.Date(treat_plan_end) >= x$monday) | is.na(treat_plan_end)
    )
    
    # Count all the plans that started before/during the week and ended during/after
    weekplans <- subset(xplans, start_bod & end_doa)
    
    # Summarize and return
    data.frame(n_active = sum(weekplans$treat_plan_type %in% "Active"),
               n_ltbi = sum(weekplans$treat_plan_type %in% "LTBI"))

})


# Reshape it for plotting
weeklytx.m <- melt(weeklytx, id.var = "monday", measure.var = c("n_active", "n_ltbi"))

# Rename the variables
weeklytx.m$variable <- as.character(weeklytx.m$variable)
weeklytx.m$variable[weeklytx.m$variable %in% "n_active"] <- "Active"
weeklytx.m$variable[weeklytx.m$variable %in% "n_ltbi"] <- "Latent"


# Plot it
ggplot(weeklytx.m, aes(x = monday, y = value, group = variable, color = variable)) +
    geom_point(size = point_size) +
    geom_line(aes(linetype = variable), size = line_size) +
    expand_limits(y = 0) +
    scale_x_date(breaks = date_breaks("1 week")) +
    scale_color_manual("Plan Type", values = colorwheel[1:3]) +
    scale_linetype_discrete("Plan Type") +
    labs(x = "Date", 
         y = "Number of patients on treatment",
         title = "Patients Being Treated for Active and Latent Tuberculosis") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, hjust = 1.30, vjust = 1.3))



```


```{r, results='asis'}

# Pretty-print
names(weeklytx) <- c("Week", "Pts on Active Tx", "Pts on Latent Tx")

dfprintif(weeklytx)

```


## Patients currently receiving treatment for Active TB
```{r active_tx}

# Query open active treatment plans
act_plans <- sqlQuery(plus, "
                      
    SELECT t.person_id, d.last_name, d.first_name, 
           t.treat_plan, t.treat_plan_date, t.plan_author,
           t.latest_tx, t.n_tx_completed
    FROM Tx_Plan_View t LEFT OUTER JOIN Demos_view d
    ON t.person_id = d.person_id
    WHERE treat_plan_type = 'Active'
        AND plan_status = 'Open'
        AND author_affiliation = 'Denver Metro TB Clinic'
        AND treat_plan_date > #2009-01-01#
                      
")


dfprintif(arrange(act_plans, plan_author, latest_tx),
          norowmsg = "It appears that there are no patients receiving treatment for active TB at this time.")


```



## Upcoming DOTs (Next Three Days)
```{r upcoming_dots}

# Query DOTs coming up in the next week
dots <- sqlQuery(plus, "
                 
    SELECT t.person_id, d.last_name, d.first_name, t.treatment_date
    FROM Drug_Treatment t LEFT OUTER JOIN Demos_View d
    ON t.person_id = d.person_id
    WHERE dispense_type IN ('DOT', 'DOPT')
        AND t.treatment_date >= Date()
        AND t.treatment_date <= Date() + 3
                 
")


dfprintif(arrange(dots, treatment_date, person_id),
          norowmsg = "It appears that there are no upcoming DOTs at this time.")

```


## Upcoming LTBI Pickups (Next Seven Days)
```{r upcoming_pickups}

# Query pickups coming up in the next week
pickups <- sqlQuery(plus, "
                 
    SELECT t.person_id, d.last_name, d.first_name, t.treatment_date
    FROM Drug_Treatment t LEFT OUTER JOIN Demos_View d
    ON t.person_id = d.person_id
    WHERE dispense_type = 'Pickup'
        AND t.treatment_date >= Date()
        AND t.treatment_date <= Date() + 7
                 
")


dfprintif(arrange(pickups, treatment_date),
          norowmsg = "It appears that there are no upcoming LTBI pickups at this time.")

```




</div>





<div id="contacts">


## Contacts Screened

```{r contactplot}

# Query all contacts and their test dates
contact.tsts <- sqlQuery(plus, paste("
                         
    SELECT c.contact_person_id AS person_id, 
           t.date_given AS test_date, 
           t.result
    FROM Contact c LEFT OUTER JOIN TST_View t
    ON c.contact_person_id = t.person_id
    WHERE t.result IN ('Negative', 'Positive')
        AND date_given BETWEEN #",
    start_date,
    "# AND #",
    stop_date,
    "#",
    sep = "")
                         
)


contact.qfts <- sqlQuery(plus, "
                         
    SELECT c.contact_person_id AS person_id, 
           q.collection_date,
           q.result
    FROM Contact c LEFT OUTER JOIN QFT_View q
    ON c.contact_person_id = q.person_id
    WHERE q.result IN ('Negative', 'Positive', 'Indeterminate')
                         
")


# Convert test dates to Dates
contact.tsts$test_date <- as.Date(contact.tsts$test_date)
contact.qfts$test_date <- as.Date(contact.qfts$collection_date, format = "%m/%d/%Y")

# Subset to QFTs in the period
contact.qfts <- subset(contact.qfts,
                       test_date >= start_date & test_date <= stop_date)

# Add test indicators before stacking
contact.tsts$test <- "TST"
contact.qfts$test <- "QFT"


# Stack 'em
contact.tests <- rbind(contact.tsts[ , c("person_id", "test_date", "test", "result")],
                       contact.qfts[ , c("person_id", "test_date", "test", "result")])


# Aggregate by test week
contact.tests$test_week <- format(contact.tests$test_date, "%Y-%W")

contact.testagg <- ddply(contact.tests, .var = "test_week", .fun = summarise,
                         freq = length(test_date),
                         week_date = get_week_start(as.Date(max(test_date, na.rm = TRUE)))
)


# Plot it
ggplot(contact.testagg, aes(x = week_date, y = freq)) +
    geom_point(size = point_size, color = colorwheel[1]) +
    geom_line(size = line_size, color = colorwheel[1]) +
    expand_limits(y = 0) +
    scale_x_date(breaks = date_breaks("1 week")) +
    labs(x = "Test week", 
         y = "Number of contacts tested",
         title = "Screenings of Contacts to Active TB Cases") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, hjust = 1.30, vjust = 1.3))



```


```{r contacttable, results = 'asis'}


# Pretty-print
names(contact.testagg) <- c("test_week", "Number of Contacts Tested", "Week")

dfprintif(contact.testagg[ , c("Week", "Number of Contacts Tested")])

```




## Current Contact Investigations



```{r investigations}

# Summarize the open contact investigations
investigations <- sqlQuery(plus, "
                           
    SELECT c.person_id, d.last_name, d.first_name,
           c.date_opened, c.reason_for, c.priority, c.lead_investigator
    FROM Contact_Investigation c LEFT OUTER JOIN Demos_View d
    ON c.person_id = d.person_id
    WHERE c.status = 'Open'
                           
")

# Summarize the contacts themselves
contacts <- sqlQuery(plus, "
  
    SELECT source_person_id, contact_person_id, priority, contact_status
    FROM Contact
                     
")

contact_sum <- ddply(contacts, .var = "source_person_id", .fun = summarise,
                     n_contacts = length(contact_person_id),
                     n_open = sum(contact_status %in% 1, na.rm = TRUE),
                     priority_one = sum(priority %in% 3, na.rm = TRUE),
                     priority_two = sum(priority %in% 2, na.rm = TRUE),
                     priority_three = sum(priority %in% 1, na.rm = TRUE)
)


# Merge contact info onto investigation info
invest_sum <- merge(x = investigations,
                    y = contact_sum,
                    by.x = "person_id",
                    by.y = "source_person_id",
                    all.x = TRUE)


# Convert it all to character for better printing
invest_print <- arrange(data.frame(lapply(invest_sum, as.character)), 
                        lead_investigator, date_opened)

dfprintif(invest_print,
          norowmsg = "It appears that there are no open contact investigations at this time.")

```




  </div>


</div>






```{r cleanup}

# Close the database connection
odbcClose(plus)


```




